---
title: "Comment on Pruitt et al (2012) 'Behavioral Types of Predator and Prey Jointly Determine Prey Survival: Potential Implications for the Maintenance of Within-Species Behavioral Variation'"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    keep_md: true
    toc: false

bibliography: bib/bib.bib
csl: bib/the-american-naturalist.csl   
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, autodep = TRUE)
library(here)
library(tidyverse)
library(janitor)
library(ggalt)
library(patchwork)
library(simba)  
library(ggVennDiagram)
library(car)
library(broom)
library(kableExtra)
source(here("R", "sim.R")) # modified version of simba::sim to return both upper and lower triangle of distance matrix

set.seed(314)
theme_set(theme_classic())
```


```{r download, include = FALSE}
#create data directory if needed
if(!fs::dir_exists(here("data"))){
  fs::dir_create(here("data"))
}

target_file <-  "Pruitt et al. Dryad File_AmNat2012.xls"
#download data if needed
if(!fs::file_exists(here("data", target_file))){
  print("Downloading file")
  filelocation <- rdryad::dryad_download("10.5061/dryad.190pk253")
  fs::file_move(
    path = filelocation[[1]][1], 
    new_path = here("data", target_file)
  ) 
}
```


```{r import, message = FALSE}
raw_data <- readxl::read_excel(
  path = here("data", target_file), 
  sheet = "raw data") %>% 
  clean_names() %>% 
  filter(!is.na(mesocosm)) %>% #remove mean & SD calculations
  rename(size = n_max)

shell_size_fear <- readxl::read_excel(
  path = here("data", target_file), 
  sheet = "shell size by fear") %>% 
  clean_names()

#number of permutations for Ruggiero similarity null 
n_ruggiero_rep <- 1000
```

# Introduction

Several articles co-authored by Professor Pruitt have been shown to contain unexplained duplications in the raw data (e.g., @Laskowski2020-retract).
This Comment reports an evaluation of whether such duplicates exist in the data for @Pruitt2012.

```{r n_snails}
#range of number of snails per mesocosm
n_snails <- raw_data %>% 
  count(mesocosm) %>% 
  summarise(min = min(n), mx = max(n), low  = sum(n < 37), n_meso = n())
```

```{r backtransform-sizes}
#transform back to mm from z scores

snail_sd <- 3.12
snail_mean <- 17.46

raw_data <- raw_data %>% 
  mutate(size = size * snail_sd + snail_mean) %>% 
  assertr::verify(near(size, round(size, 2))) %>% # check all are 2 decimal place
  mutate(size = round(size, 2)) # remove rounding errors
```


```{r last-digit}
last_digit <- raw_data %>% 
  mutate(last = round(size * 100) %% 10) %>% 
  count(last) %>% 
  mutate(p = n/sum(n) * 100) # huge excess of 1

last_digit_out <- list()
last_digit_out$percent_last1 <- last_digit %>% 
  filter(last == 1) %>%
  pull(p) %>% 
  round(1)

last_digit_out$shell_size_fear_last <- shell_size_fear %>% 
   mutate(last = (round(shell_diameter * 100) %% 10)) %>% 
  count(last) %>% 
  mutate(p = n/sum(n) * 100) # huge excess of 1
```

# Data description

Partial raw data are archived as an Excel file [@Pruitt2011_data].
The sheet `raw data` includes data for activity, size (as `N Max`), anti-predator response, and survivorship for `r nrow(raw_data)` snails in `r max(raw_data$mesocosm)` mesocosms across three replicates (A-C).
Size and anti-predator response are presented as _z_-scores. 
Size can be back-transformed to millimetres with the mean and standard deviation reported in the article.
After back-transformation, a large fraction of the measurements (`r last_digit_out$percent_last1`%) end in a one, perhaps suggesting that the reported mean is 0.01 mm too large and that most snails were measured to one decimal place, and the rest to two. 
This is supported by a second Excel sheet (`shell size by fear`) which reports snail sizes in millimetres (without metadata) has a similarly large fraction of measurements with a terminal zero.
Mesocosms have between `r n_snails$min` and `r n_snails$mx` snails, whereas the article reports that 37--46 snails are used in each mesocosm: `r n_snails$low` mesocosms have fewer snails than the reported minimum.
The number of snails per mesocosm is varied in replicates A and B, but mostly consistent in replicate C.
Within each mesocosm the data are sorted by survival and partially sorted by snail size.

```{r n-snail-plot, fig.cap = "Number of snails per mesocosm in the three replicates apparently measured to one (dark purple) or two (light green) decimal places.", fig.height = 3, fig.width = 6, include=FALSE}
raw_data %>% 
  mutate(
    last = round(size * 100) %% 10,
    last1 = factor(last == 1, labels = c("2 D.P.", "1 D.P."))
    ) %>% 
  ggplot(aes(x = factor(mesocosm), fill = last1)) + 
  geom_bar() + 
  facet_wrap(~replicate, scales = "free_y") +
  scale_fill_viridis_d(end = 0.8, limits = c("1 D.P.", "2 D.P.")) + 
  coord_flip() +
  labs(x = "Mesocosm", y = "Count", fill = "Apparent resolution") +
  theme(legend.position = "none") 
```

```{r raw-data-plot, fig.cap = "Size of snails that died (blue) or survived (orange) against row number from three mesocosm.", fig.height = 3.5, fig.width = 4, include = FALSE}
#example of raw data
raw_data %>% 
  group_by(mesocosm) %>% 
  mutate(row = 1:n()) %>%
  filter(mesocosm %in% c(1:3)) %>% 
  ggplot(aes(x = row, y = size, colour = factor(survival_1_0))) +
  geom_point(show.legend = FALSE) +
  scale_colour_viridis_d(option = "C", end = 0.8) +
  facet_wrap(~mesocosm, ncol = 1) +
  labs(x = "Row", y = "Size mm", colour = "Survival")
```

# Methods

Because the data have been sorted, tests for duplicate sequences such as sequenceSniffer [@Rutten2020] have a reduced utility because sequences may have been disrupted by sorting, or false positive duplicate sequences generated.

Instead, I test whether the overlap between measurements for the different mesocosms is higher than expected. 
With a slightly modified version of the `simba` package [@simba] in R [@R], I calculate the Ruggiero similarity [@Ruggiero1998] between mesocosms. 
The Ruggiero similarity ($a/(a + c)$ where $a$ is the number of shared attributes and $c$ is the number of attributes found only in the second sample) is one when all members of the second sample are in the first. 
I make all repeated values distinct, so if there are repeated values in mesocosm 2, there needs to be at least as many repeats of that value in mesocosm 1 to get a similarity of one.

The expected distribution of Ruggiero similarities is estimated by permuting the snail sizes within each replicate and recalculating the similarities `r n_ruggiero_rep` times.
This mimics the description in the article which reports that "Randomly selected, individually marked snails ...  were placed within each ... mesocosms".

```{r ruggiero}
# similarity between size measurements in different mesocosms
ruggiero <- raw_data %>% 
  group_by(mesocosm, size) %>% 
  mutate(size2 = paste(size, 1:n(), sep = "_")) %>% #make unique within each mesocosm
  ungroup() %>% 
  count(mesocosm, size2) %>% # formats data as required by simba: three columns representing plots, species and occurrence information 
  sim(method = "ruggiero", listout = FALSE, listin = TRUE) 

#convert distance matrix to long tibble
 ruggiero2 <- ruggiero %>% 
    as.data.frame() %>% 
    rownames_to_column(var = "xid") %>% 
    pivot_longer(-xid, names_to = "yid", values_to = "rug") %>% 
    filter(xid != yid) %>%  # remove diagonal
    mutate(across(everything(), as.numeric))
 
# Null model. Permutation of sizes within replicates
# find similarity of mesocosms with permuted data 
random_ruggiero <- rerun(n_ruggiero_rep, {
  ruggiero <- raw_data %>%
    group_by(replicate) %>% 
    mutate(size = sample(size)) %>% #randomise
    group_by(mesocosm, size) %>% 
    mutate(size2 = paste(size, 1:n(), sep = "_")) %>% 
    ungroup() %>% 
    count(mesocosm, size2) %>% 
    sim(method = "ruggiero", listout = FALSE, listin = TRUE) 

 ruggiero %>% 
    as.data.frame() %>% 
    rownames_to_column(var = "xid") %>% 
    pivot_longer(-xid, names_to = "yid", values_to = "rug") %>% 
    filter(xid != yid) %>%  # remove diagonal
    mutate(across(everything(), as.numeric))
})

#find max of random trials
rugg_rand_max <- random_ruggiero %>% 
  set_names(1:length(.)) %>% 
  bind_rows(.id = "n") %>% 
  group_by(n) %>% 
  summarise(mx = max(rug), .groups = "drop") 
```

# Results

```{r replicate-C}
snail_meta <- raw_data %>% distinct(replicate, mesocosm)

ruggiero3 <- ruggiero2 %>%
  left_join(snail_meta, by = c(xid = "mesocosm")) %>% 
  mutate(xid = factor(xid), xid = fct_reorder(xid, as.numeric(factor(replicate)))) %>% 
  rename(replicate1 = replicate) %>% 
  left_join(snail_meta, by = c(yid = "mesocosm")) %>% 
  mutate(yid = factor(yid), yid = fct_reorder(yid, as.numeric(factor(replicate)))) %>% 
  rename(replicate2 = replicate) %>% 
  mutate(rug_class = case_when(rug  == 1 ~ "1",
                          rug > 0.9 ~ "> 0.9",
                          rug > 0.8 ~ "> 0.8",
                          rug > 0.7 ~ "> 0.7",
                          TRUE ~ "< 0.7")) 

C_max <- ruggiero3 %>% 
  group_by(xid) %>% 
  filter(replicate1 == "C") %>% 
  slice(which.max(rug)) %>% 
  ungroup() 
C_min <- C_max %>% 
  summarise(m = min(rug))
```


```{r ruggiero-plots, fig.cap = glue::glue("For the snail size data, A) distribution of Ruggiero similiarities between the mesocosms in the archived data, B) maximum Ruggiero similiarity between mesocosms for each of {n_ruggiero_rep} permutations, C) Ruggiero similiarities between pairs of mesocosms (red lines separate annotated replicates)."), fig.width=8, fig.height=9}

labs <- seq(0, 1, 0.1)
labs[c(FALSE, TRUE)] <- ""

#raw data histogram
rugg_size_hist <- ruggiero2 %>% 
  ggplot(aes(x = rug)) + 
  geom_histogram(bins = 30, boundary = 0) +
  scale_x_continuous(limits = c(0, 1), expand = c(0.01, 0), breaks = seq(0, 1, 0.1), labels = labs) +
  annotate(geom = "segment", 
           x = max(ruggiero2$rug), 
           xend = max(ruggiero2$rug), 
           y = 220,
           yend = 0, 
           colour = "red", 
           arrow = arrow(length = unit(0.03, "npc"))) +
  annotate(geom = "text", label = "Maximum", x = max(ruggiero2$rug), y = 225, vjust = 0, hjust = 0.9, colour = "red") +
  labs(x = "Ruggiero similarity in raw data")
 
#random max similarity
rugg_rand_hist <- rugg_rand_max %>% 
  ggplot(aes(x = mx)) +
  geom_histogram(bins = 30, boundary = 0) +
  annotate(geom = "segment", 
           x = max(rugg_rand_max$mx), 
           xend = max(rugg_rand_max$mx), 
           y = 200,
           yend = 0, 
           colour = "red", 
           arrow = arrow(length = unit(0.03, "npc"))) +
  annotate(geom = "text", label = "Maximum", x = max(rugg_rand_max$mx), y = 205, vjust = 0, colour = "red") +
  scale_x_continuous(limits = c(0, 1), expand = c(0.01, 0), breaks = seq(0, 1, 0.1), labels = labs) +
  labs(x = "Maximum Ruggiero similarity in permutated data")

#raster

rugg_raster <- ruggiero3 %>% 
  ggplot(aes(x = xid, y = yid, fill = rug_class)) + 
    geom_raster() +
    scale_fill_viridis_d()  + 
    geom_vline(xintercept = c(18.5, 36.5), colour = "red") +
    geom_hline(yintercept = c(18.5, 36.5), colour = "red") +
    labs(x = "Mesocosm 1", y = "Mesocosm 2", fill = "Ruggiero\nsimilarity") +
    theme(panel.grid = element_blank(), 
          axis.text.x = element_text(angle = 90, vjust = 0.5)) +
   annotate_textp("A - A", x = 1/6, y = 0.3, hjust = 0.5, vjust = 0, color = "white", box_just = 0.5) + 
  annotate_textp("A - B", x = 1/6, y = 0.64, hjust = 0.5, vjust = 0, color = "white", box_just = 0.5) + 
  annotate_textp("A - C", x = 1/6, y = 0.98, hjust = 0.5, vjust = 0, color = "white", box_just = 0.5) +
   annotate_textp("B - A", x = 1/2, y = 0.3, hjust = 0.5, vjust = 0, color = "white", box_just = 0.5) + 
  annotate_textp("B - B", x = 1/2, y = 0.64, hjust = 0.5, vjust = 0, color = "white", box_just = 0.5) + 
  annotate_textp("B - C", x = 1/2, y = 0.98, hjust = 0.5, vjust = 0, color = "white", box_just = 0.5) +
  annotate_textp("C - A", x = 5/6, y = 0.3, hjust = 0.5, vjust = 0, color = "white", box_just = 0.5) + 
  annotate_textp("C - B", x = 5/6, y = 0.64, hjust = 0.5, vjust = 0, color = "white", box_just = 0.5) + 
  annotate_textp("C - C", x = 5/6, y = 0.98, hjust = 0.5, vjust = 0, color = "white", box_just = 0.5) 

#compile plots with patchwork
(rugg_size_hist / rugg_rand_hist) / rugg_raster +
  patchwork::plot_annotation(tag_levels = "A") + 
  plot_layout(heights = c(0.2, 0.2, 0.6))
```

Some pairs of mesocosms have much higher Ruggiero similarities than expected from a random permutation of the data (Fig. \@ref(fig:ruggiero-plots)).
For example, `r sum(ruggiero2$rug == 1)` pairs of mesocosms have a Ruggiero similarity of 1 (Fig. \@ref(fig:ruggiero-plots)A), that is the measurements in the smaller mesocosm are a subset of those in the larger mesocosm.
A further `r sum(ruggiero2$rug > 0.8 & ruggiero2$rug < 1)` pairs of mesocosms have a Ruggiero similarity above 0.8 (Fig. \@ref(fig:ruggiero-plots)A).
In contrast, the 99^th^ percentile of the maximum similarity found in `r n_ruggiero_rep` permutations is `r quantile(rugg_rand_max$mx, prob = 0.99) %>% round(2)` and the highest maximum observed was `r max(rugg_rand_max$mx) %>% round(2)` (Fig. \@ref(fig:ruggiero-plots)B).
Similarities between mesocosms above 0.8 are therefore expected to be very rare, and datasets with many such similarities should be exceedingly rare.


There are a few pairs of mesocosms with high similarities within replicates, but most of the pairs are when mesocsoms from replicates A and B are subsets of the larger replicate C mesocosms. 
All of the mesocosm in replicate C have a similarity of at least `r round(C_min$m, 3)` with mesocosms in replicates A or B.

```{r coverage}
last <- raw_data %>% 
  select(replicate, mesocosm, size) %>% 
  mutate(dp = if_else(round(size * 100, 2) %% 10 == 1, "one", "two"))

coverage <- last %>% 
  distinct(dp, size) %>%
  mutate(int = floor(size)) %>% 
  group_by(dp) %>% 
  count(int) %>% #ignores zero counts
  summarise(m = mean(n)) %>% 
  mutate(p = if_else(dp == "one", m / 10, m / 90))
```

There are no size measurements unique to replicate C (Fig. \@ref(fig:unique-to-C)), even though this replicate has the most snails, and, at least for the data apparently measured to two decimal places, only a small proportion (< 10%) of possible sizes have been used.

```{r unique-to-C, fig.cap = "Venn diagram of the size measurements by replicate", fig.height = 3.5, fig.width = 4.5}
raw_data %>% 
  distinct(replicate, size) %>% #unnecessary as Venn diagram uses sets
  group_by(replicate) %>% 
  group_split() %>% 
  set_names(c("A", "B", "C")) %>% 
  map(pull, size) %>% 
  ggVennDiagram()
```

High similarities were not detected in either the activity or anti-predator response data.



## Overlap between replicate C and the other replicates

This section explores the high similarity between mesocosms in replicate C and the other two replicates in more detail.
Figure \@ref(fig:replicates-A-B-vs-C) shows for each snail in replicate C (one column per mesocosm) whether there is a snail the same size in each mesocosm from replicates A and B (rows, arranged by mesocosm number).
Some measurements are expected to in common between the pairs of mesocosms, but long sequences are not.

Mesocosm 3 includes all measurements from mesocosm 4; measurements not in mesocosm 4 are in mesocosm 2. 
Almost all the measurement from mesocosm 6 overlap with measurements from mesocosm 5.
Further across, about half of mesocosm 12 overlaps with mesocosm 8, and half with mesocosm 10, which in turn overlaps the first part of mesocosm 15.
Similar overlaps of whole or part mesocosms appear for most or all of replicate C.

```{r replicates-A-B-vs-C, fig.height = 9, fig.width = 9.5, fig.cap = "Measurements in common (purple) between mesocosms in replicate C (columns) and mesocosms in replicates A and B (rows). Axes are size against snail row number within replicate C.", fig.height = 8, fig.width = 8}
#extract replicate C mesocosms
replicate_C <- raw_data %>% 
  filter(replicate == "C") %>% 
  select(mesocosmC = mesocosm, size) %>% 
  mutate(n = 1:n()) #sequential numbers


raw_data %>% 
  filter(replicate != "C") %>% #get other replicates
  select(replicate, mesocosm, size) %>% 
  group_by(replicate, mesocosm) %>% 
  nest() %>% 
  mutate(df = map(data, # for each mesocosm in A/B find matching measurements in replicate C
                  ~left_join(
                    replicate_C, 
                    .x %>% mutate(one = 1), #dummy column to show match made
                    by = "size"))) %>% 
  select(-data) %>% 
  unnest(df) %>% 
  ggplot(aes(x = n, y = size, colour = is.na(one))) +
  geom_point(show.legend = FALSE, size = 1) +
  scale_colour_viridis_d(end = 0.9) +
  facet_grid(mesocosm ~ mesocosmC, scales = "free_x") +
  scale_y_continuous(breaks = c(15, 25)) +
  labs(y = "Size mm", x = "Row") +
  theme(panel.spacing = unit(0.4, "mm"), 
        strip.text.y = element_text(angle = 0), 
        axis.text.x = element_blank()
        )
```

## Authors' Response

In their response to the first version of this comment, the authors wrote:

> snails with identical shell diameters were deliberately split between mesocosms, thus strong overlap between sets of mesocosms set up at the same time is expected. 

This sets up two testable predictions: that there should be fewer snails with identical sizes in each mesocosm than expected under a null model of random assignment; and that pairs of mesocosms with high similarities should tend to occur in the same replicate.

```{r duplicates-test}
dups <- raw_data %>% 
  group_by(replicate, mesocosm) %>% 
  count(size) %>% 
  filter(n > 1) %>% 
  summarise(mx = max(n), sum = sum(n), number = n(), dup = sum - number, .groups = "drop_last")

dups_summary <- dups %>% 
  summarise(across(c(mx, sum, number, dup), mean), .groups = "drop")

dups_rand <- rerun(n_ruggiero_rep, 
  raw_data %>%
     mutate(last1 = round(size * 100) %% 10 == 1) %>% 
  group_by(replicate, last1) %>% # make sure distribution of 1 & 2 dp values is maintained
  mutate(size = sample(size)) %>% 
  group_by(replicate, mesocosm) %>% 
  count(size) %>% 
  filter(n > 1) %>% 
  summarise(sum = sum(n), number = n(), dup = sum - number, .groups = "drop_last") %>% 
  summarise(across(c(sum, number, dup), mean), .groups = "drop")
  ) %>% 
  bind_rows(.id = "run")
```


```{r duplicates-test-plot, fig.cap = glue::glue("Mean number of repeated shell sizes in each mesocosm per replicate. Observed number shown by red dots, distribution of repeats in {n_ruggiero_rep} permutations shown by violins."), fig.height = 3.5, fig.width = 4}
ggplot(dups_rand, aes(x = replicate, y = dup)) + 
  geom_violin(draw_quantiles = c(0.25, 0.75), alpha = 0.3, fill = "grey70") +
  geom_point(data = dups_summary, colour =  "red", size = 3, position = position_dodge(width = 1)) +
  labs(x = "Replicate", y = "Mean no. repeated sizes per mesocosm")
```


Neither prediction is supported by the the data:
repeated sizes are not less common than expected (Fig. \@ref(fig:duplicates-test-plot);
and most pairs of mesocosms with high similarity are between rather than within replicates (Fig. \@ref(fig:replicates-A-B-vs-C)).

# Conclusions

```{r how-unlikely}
#pbinom
#lower triangle
rug_sum <- ruggiero3 %>% 
  filter(paste(replicate1, xid) > paste(replicate2, yid)) %>% 
  summarise(n = n(), high = sum(rug > 0.8))


rug_prob <- pbinom(q = rug_sum$high - 1,
       size = rug_sum$n, 
       prob = 1/(rug_sum$n * n_ruggiero_rep), lower.tail = FALSE)
```

There are unexpectedly high similarities between size measurements in the different mesocosms.
An important question is how unlikely are the observed patterns in the data to occur by chance.
With the result presented above, it is possible to set an upper bound on this probability.
Considering only the lower triangle of the distance matrix (as the upper triangle is not independent), `r n_ruggiero_rep` $\times$ `r rug_sum$n` pairs of mesocosms were tested to yield the maximum similarity of `r max(rugg_rand_max$mx) %>% round(2)`, so pairs of mesocosms with a similarity > 0.8 must have a probability of occurrence less than `r (1/(n_ruggiero_rep * rug_sum$n)) %>% signif(2)`. 
The probability of getting `r rug_sum$high` or more such pairs is `r (rug_prob * 2 * 3) %>% signif(2)` after allowing for the choice of the upper or lower triangle and size rather than the other response variables.

Snail size is an important variable in @Pruitt2012.
This comment shows that it is very unlikely to have been collected using the methods reported in the original article.
If it cannot be relied upon, parts of the results become questionable.

# Reproducibility

All analyses in this comment can be reproduced with an rmarkdown document archived at https://github.com/Sorbus-torminalis/Pruitt_2012_Am_Nat

# References {-}

<div id="refs"></div>

# Appendix A {-}

If the snails are randomly allocated to mesocosms within each replicate, one would expect the means and variances to be similar.
This can be tested with ANOVAs and Levene's tests.
All three ANOVA (Table \@ref(tab:lm-levene)), and two of the Levene's tests are highly significant, suggesting that the randomisation was not very effective. 
Figure \@ref(fig:mesocosm-size-plots) shows that some mesocosms in each replicate have smaller sized snails that expected 

```{r lm-levene}
#differences in means
mod_A <- raw_data %>% 
  filter(replicate == "A") %>% 
  mutate(mesocosm = factor(mesocosm)) %>% 
  lm(size ~ mesocosm, data = .)

mod_B <- raw_data %>% 
  filter(replicate == "B") %>% 
  mutate(mesocosm = factor(mesocosm)) %>% 
  lm(size ~ mesocosm, data = .)

mod_C <- raw_data %>% 
  filter(replicate == "C") %>% 
  mutate(mesocosm = factor(mesocosm)) %>% 
  lm(size ~ mesocosm, data = .)

anovas <- list(A = mod_A, B = mod_B, C = mod_C) %>% 
  map_dfr(broom::glance, .id = "Replicate") %>% 
  select(Replicate, df, `residual df` = df.residual, `F` = statistic, `p value` = p.value) %>% 
  mutate(`p value` = format(`p value`, digits = 2)) 


# Levenes tests homogeneity_variance
Lev_A <- leveneTest(size ~ factor(mesocosm), data = raw_data %>% filter(replicate == "A"), center = "median")
Lev_B <- leveneTest(size ~ factor(mesocosm), data = raw_data %>% filter(replicate == "B"), center = "median")
Lev_C <- leveneTest(size ~ factor(mesocosm), data = raw_data %>% filter(replicate == "C"), center = "median")

levenes <- list(A = Lev_A, B = Lev_B, C = Lev_C) %>% 
  map_dfr(broom::tidy, .id = "Replicate") %>% 
  select(Replicate, df, `residual df` = df.residual, `F` = statistic, `p value` = p.value) %>% 
  mutate(`p value` = format(`p value`, digits = 2))

left_join(anovas, levenes, by = c("Replicate", "df", "residual df"), suffix = c("", " ")) %>% 
  kbl(booktabs = TRUE, digits = 2, caption = "ANOVA and Levene's tests of snail sizes in the different mesocosms in each replicate.") %>%
  kable_styling(latex_options = "striped") %>% 
  add_header_above(c(" " = 3, "ANOVA" = 2, "Levene's test" = 2))
```


```{r mesocosm-size-plots, fig.cap = "Snail size for each mesocosm in each replicate. Horizontal lines are the mean for each replicate.", fig.width = 6, fig.height = 3.5}
raw_data %>% 
  mutate(mesocosm = factor(mesocosm)) %>% 
  ggplot(aes(x = mesocosm, y = size, fill = replicate)) +
  geom_boxplot(show.legend = FALSE) +
  geom_hline(aes(yintercept = y), raw_data %>% 
               group_by(replicate) %>% 
               summarise(y = mean(size), .groups = "drop")) +
  scale_fill_brewer(palette = "Set1") +
  labs(x = "Mesocosm", y = "Size mm") +
  facet_wrap(~ replicate, scales = "free_x") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```
```{r cache=FALSE}
#don't make remaining plots
knitr::knit_exit()
```


# Appendix B {-}

## Within replicate A

```{r replicateA}
## within A
raw_data %>% 
  filter(replicate == "A") %>% 
  select(replicate, mesocosm, size) %>% 
  group_by(replicate, mesocosm) %>% 
  nest() %>% 
  mutate(df = map(data, ~left_join(raw_data %>% mutate(n = 1:nrow(.)) %>% 
  filter(replicate == "A") %>% 
  select(mesocosmA = mesocosm, size, n), .x %>% mutate(one = 1), by = "size"))) %>% 
  select(-data) %>% 
  unnest(df) %>% 
  filter(mesocosm != mesocosmA) %>% 
  ggplot(aes(x = n, y = size, colour = is.na(one))) +
  geom_point(show.legend = FALSE, size = 1) +
   scale_colour_viridis_d(end = 0.9) +
  facet_grid(mesocosm ~ mesocosmA, scales = "free_x") +
  scale_y_continuous(breaks = c(15, 25)) +
  labs(y = "Size mm", x = "Row") +
  theme(panel.spacing = unit(0.4, "mm"), 
        strip.text.y = element_text(angle = 0), 
        axis.text.x = element_blank()
        )
```

## Within replicate B {-}

```{r replicateB}
## within B
raw_data %>% 
  filter(replicate == "B") %>% 
  select(replicate, mesocosm, size) %>% 
  group_by(replicate, mesocosm) %>% 
  nest() %>% 
  mutate(df = map(data, ~left_join(raw_data %>% mutate(n = 1:nrow(.)) %>% 
  filter(replicate == "B") %>% 
  select(mesocosmB = mesocosm, size, n), .x %>% mutate(one = 1), by = "size"))) %>% 
  select(-data) %>% 
  unnest(df) %>% 
  filter(mesocosm != mesocosmB) %>% 
  ggplot(aes(x = n, y = size, colour = is.na(one))) +
  geom_point(show.legend = FALSE, size = 1) +
  scale_colour_viridis_d(end = 0.9) +
  facet_grid(mesocosm ~ mesocosmB, scales = "free_x") + 
  scale_y_continuous(breaks = c(15, 25)) +
  labs(y = "Size mm", x = "Row") +
  theme(panel.spacing = unit(0.4, "mm"), 
        strip.text.y = element_text(angle = 0), 
        axis.text.x = element_blank()
        )
```


## Replicate A vs replicate B {-}

```{r replicates-A-vs-B}
## A vs B
raw_data %>% 
  filter(replicate == "A") %>% 
  select(replicate, mesocosm, size) %>% 
  group_by(replicate, mesocosm) %>% 
  nest() %>% 
  mutate(df = map(data, ~left_join(raw_data %>% mutate(n = 1:nrow(.)) %>% 
  filter(replicate == "B") %>% 
  select(mesocosmB = mesocosm, size, n), .x %>% mutate(one = 1), by = "size"))) %>% 
  select(-data) %>% 
  unnest(df) %>% 
  filter(mesocosm != mesocosmB) %>% 
  ggplot(aes(x = n, y = size, colour = is.na(one))) +
  geom_point(show.legend = FALSE, size = 1) +
  scale_colour_viridis_d(end = 0.9) +
  facet_grid(mesocosm ~ mesocosmB, scales = "free_x") +
  scale_y_continuous(breaks = c(15, 25)) +
  labs(y = "Size mm", x = "Row") +
  theme(panel.spacing = unit(0.4, "mm"), 
        strip.text.y = element_text(angle = 0), 
        axis.text.x = element_blank()
        )
```
